## Markdown file with R code snippets

## Install packages
```{r}
install.packages("tidyverse")
```

## Load Libararies
```{r}
library(tidyverse)
```

## Create Data Frames
This creates data frames directly in R
First, a vector of names
```{r}
names <- c("Bob", "Mike", "Amy", "Jesper")
```

Then a vector of ages
```{r}
age <- c(16, 18, 26, 23)
```

Combine those vectors to create a simple data frame
```{r}
people <- data.frame(names, age)
```

## Inspect the Data Frame
Preview the data using the head or tail command.
In this cases, the data set is small enough that either one will displace all of the data in the frame
```{r}
head(people)
```

```{r}
tail(people)
```

Inspect the internal structure of the data frame using the str command
```{r}
str(people)
```

Inspect the columns and a preview of the data using the glimpse command
```{r}
glimpse(people)
```

Inspect the columns without looking at the data itself using the colnames command
```{r}
colnames(people)
```

## Make Alterations
Use the mutate command to see the ages of all individuals in 20 years
```{r}
mutate(people, age_in_20 = age + 20)
```

## Repeat for New Data
Follow the steps above to create a data frame about fruit

First create the vectors
```{r}
names <- c("Apple", "Banana", "Grapefruit", "Pear", "Kiwi", "Tomato", "Pineapple", "Dragonfruit", "Strawberry", "Grape", "Cranberry", "Rasberry")
```

```{r}
color <- c("Red", "Yellow", "Yellow", "Green", "Brown", "Red", "Brown", "Pink", "Red", "Purple", "Red", "Red")
```

```{r}
rating <- c("5", "5", "4", "1", "3", "1", "4", "3", "1", "2", "3", "3")
```

Then combine vectors into a data frame
```{r}
fruit <- data.frame(names, color, rating)
```

Preview data frame
```{r}
glimpse(fruit)
```

