## Markdown file with R code snippets

## Install packages
```{r install packages}
install.packages("tidyverse")
install.packages("skimr")
install.packages("janitor")
```

## Load Libararies
```{r load packages}
library(tidyverse)
library(skimr)
library(janitor)
```

## Create Data Frames
This creates data frames directly in R
First, a vector of names
```{r}
names <- c("Bob", "Mike", "Amy", "Jesper")
```

Then a vector of ages
```{r}
age <- c(16, 18, 26, 23)
```

Combine those vectors to create a simple data frame
```{r}
people <- data.frame(names, age)
```

## Inspect the Data Frame
Preview the data using the head or tail command.
In this cases, the data set is small enough that either one will displace all of the data in the frame
```{r}
head(people)
```

```{r}
tail(people)
```

Inspect the internal structure of the data frame using the str command
```{r}
str(people)
```

Inspect the columns and a preview of the data using the glimpse command
```{r}
glimpse(people)
```

Inspect the columns without looking at the data itself using the colnames command
```{r}
colnames(people)
```

## Make Alterations
Use the mutate command to see the ages of all individuals in 20 years
```{r}
mutate(people, age_in_20 = age + 20)
```

## Repeat for New Data
Follow the steps above to create a data frame about fruit

First create the vectors
```{r}
names <- c("Apple", "Banana", "Grapefruit", "Pear", "Kiwi", "Tomato", "Pineapple", "Dragonfruit", "Strawberry", "Grape", "Cranberry", "Rasberry")
```

```{r}
color <- c("Red", "Yellow", "Yellow", "Green", "Brown", "Red", "Brown", "Pink", "Red", "Purple", "Red", "Red")
```

```{r}
rating <- c("5", "5", "4", "1", "3", "1", "4", "3", "1", "2", "3", "3")
```

Then combine vectors into a data frame
```{r}
fruit <- data.frame(names, color, rating)
```

Preview data frame
```{r}
glimpse(fruit)
```


## Import Data
Import data from a csv file
```{r}
bookings_df <- read_csv("hotel_bookings.csv")
```

## Inspect and Clean Data
Preview data
```{r}
head(bookings_df)
```

Create a new data frame based on average daily rate "adr"
```{r}
new_df <- select(bookings_df, `adr`, adults)
```

Make changes to the new data frame to find the adr without affecting the original data frame
```{r}
mutate(new_df, total = `adr` / adults)
```


## Make Changes to Data Frames
Import data frame from csv file
```{r load dataset}
hotel_bookings <- read_csv("")
```

Inspect the data frame
```{r head function}
head()
```

Get a more detailed summary of data
```{r}
skim_without_charts(hotel_bookings)
```

Create a cleaner dataset by combining arrival month and year into one column
```{r}
combined_df <- hotel_bookings %>%
  select(arrival_date_year, arrival_date_month) %>% 
  unite(arrival_month_year, c("arrival_date_month", "arrival_date_year"), sep = " ")
```

Clean dataset to examine only what management is focused on
```{r}
trimmed_df <- combined_df %>% 
  select(hotel_type, is_canceled, lead_time, arrival_month_year)
```

Rename hotel_type field to make it more clear
```{r}
trimmed_df %>% 
  select(hotel, is_canceled, lead_time) %>% 
  rename(hotel_type = hotel)
```

Arrange bookings by lead time in descending order
```{r arrange function descending} 
arrange(trimmed_df, desc(lead_time))
```

Inspect data frame to check changes
```{r head function part two}
head(trimmed_df)
```

Check max and min lead times
```{r}
max(trimmed_df$lead_time)
```

```{r}
min(trimmed_df$lead_time)
```

Check average lead time for hotel bookings
```{r mean}
mean(trimmed_df$lead_time)
```

Filter data to City Hotels only
```{r filter}
hotel_bookings_city <- filter(trimmed_df, trimmed_df$hotel=="City Hotel")
```

Inspect data
```{r new dataset}
head(hotel_bookings_city)
```

Check average lead time on bookings for City Hotels only
```{r average lead time city hotels}
mean(hotel_bookings_city$lead_time)
```

Provide a summary of data by hotel including mean, min, and max lead times
```{r group and summarize}
hotel_summary <- 
  trimmed_df %>%
  group_by(hotel) %>%
  summarise(average_lead_time=mean(lead_time),
            min_lead_time=min(lead_time),
            max_lead_time=max(lead_time))
```

Inspect the new data
```{r}
head(hotel_summary)
```



